---
title: "Reddit Science Analysis"
author: "Laura Fontanills"
date: "2022-12-08"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

### Required packages

```{r}
library(readr) # create data frames from .csv files
library(tidyverse) # data cleaning
library(lubridate) # wrangle dates
library(ggplot2) # visualization
library(urltools) # wrangle urls
library(tidytext) #NLP toolkit
```

### Data scraping

Use scraper.py to generate 3 .csv files, corresponding to the top 100 posts of the last month, year, and of all-time. Each of these contain 9 variables:
- `...1` ranks the post from 0-99 (highest score to lowest score)
- `id` is a unique string of letters and numbers that identify each post. Each post only has 1 id.
- `created_unix_utc` is the time the post was created as a unix number.
- `post_url` is the url for the website the post links to. This can be from an online news outlet, journal, etc.. Each post has a unique url.
- `post_title` is the title of the post.
- `flair` is a tag attached to each post by a moderator that places that post in a category. Categories are fields of science and include biology, physics, psychology, and news.
- `score` is the number of times a post was upvoted by a user (indicating that they found the post interesting and relevant to the subreddit).
- `num_comments` is the number of unique comments a post received. This does not include edits to comments.
- `upvote_ratio` is the ratio of upvotes (approval) to downvotes (disapproval) a post received. It is formatted as a decimal from 0 to 1.

```{r}
# create and view data frame with top 100 posts from the last month
top_month <- read_csv("Top-Posts-Month.csv")
View(top_month)

# create and view data frame with top 100 posts from the last year
top_year <- read_csv("Top-Posts-Year.csv")
View(top_year)

# create view data frame with top 100 posts from all time
top_all <- read_csv("Top-Posts-All.csv")
View(top_all)
```

### Data cleaning

Data frames are tidy: one row for each post id, fields OK.

Before combining data frames I am going to add columns "all_time", "past_year", and "past_month" with boolean indicating over which time period(s) the post made the top 100. 
```{r}

# before combining data frames, add a column indicating if it was a top "month", "year", or "all" post.
top_all_2 <- top_all %>% 
  mutate(all_time = TRUE) %>% 
  mutate(past_year = FALSE) %>% 
  mutate(past_month = FALSE)

top_year_2 <- top_year %>%
  mutate(all_time = FALSE) %>% 
  mutate(past_year = TRUE) %>% 
  mutate(past_month = FALSE)

top_month_2 <- top_month %>% 
  mutate(all_time = FALSE) %>% 
  mutate(past_year = FALSE) %>% 
  mutate(past_month = TRUE)
```




### Analysis 1: everything but the titles



